Dynamic Adaptive Saturated Sliding Mode Control For Deployment of Tethered Satellite System

### 2016-10-18 manuscript.tex
20161018

H(q)\ddot {q}+C(q,\dot{q})\dot{q}+G(q) = \tau,\label{eq:dynamic model}

@@

q&= \begin{pmatrix}\lambda&\theta&\phi\end{pmatrix}^T\\
\tau&=\begin{pmatrix}\hat\tau_t&\hat\tau_\theta&\hat\tau_\phi\end{pmatrix}^T\\
H(q) &= \begin{bmatrix}1 &0&0\\0 &\lambda^2\cos^2\phi&0\\0&0&\lambda^2\end{bmatrix}\\
C(q,\dot{q}) &=\begin{bmatrix}0 &-(\lambda\dot \theta+2\lambda)\cos^2\phi&-\lambda\dot\phi\\(2\lambda+\lambda\dot \theta)\cos^2\phi&\lambda\dot \lambda\cos^2\phi-\lambda^2\dot\phi\sin\phi\cos\phi&-(\dot\theta+2)\lambda^2\sin\phi\cos\phi\\ \lambda\dot\phi&\lambda^2(\dot\theta+2)\sin\phi\cos\phi&\lambda\dot \lambda\end{bmatrix}\\
G(q) &=\begin{bmatrix}-\lambda\cos^2\phi+2\lambda+3\lambda\cos ^2\theta\cos^2\phi\\3\lambda^2\cos\theta\sin\theta\cos^2\phi\\\lambda^2+3\lambda^2\cos^2\theta\sin\phi\cos\phi\end{bmatrix}.

@@

Based on the expression above, one can find that $\dot H(q) - 2C(q,\dot{q})$ is skew-symmetric, namely,

@@

q^T(\dot H(q) - 2C(q,\dot{q}))q = 0.

@@

Before declaring the controller design, some basic precondition is prepared for proceeding derivation.  The left side of Eq.~(\ref{eq:dynamic model}) is able to be denoted by the linearly parametrization $Y\Theta$, in which $Y=Y(q,\dot q,\ddot q)$ consists of known nonlinear functions. $\Theta$ denotes unknown but constant parameters, and particularly it is a constant vector in this dimensionless form. Reconstruct the parametrization $Y\Theta$ in terms of a reference $\dot q_r$:

@@

 Y_r\Theta\equiv H(q)\ddot q_r+C(q,\dot q)\dot q_r+G(q)\label{eq:dynamic model 1}

@@

where $Y_r = Y_r(q,\dot q,\dot q_r,\ddot q_r)$. Based on results mentioned above, the open-loop error dynamics expressed by $S_r$ is able to be obtained, as follows:

@@

H(q)\dot S_r+C(q,\dot q)S_r=\tau - Y_r\Theta\label{eq:dynamic model 2}

@@

S_r=\dot q-\dot q_r.\label{eq:Sr1}

@@

We define the tracking error as $\Delta q=q-q_d$, in which $q_d$ is the reference trajectory with $q_d(t)\in C^2$. For deriving the control scheme, some properties should be given out in advance.
\begin{mypro}
Positive scalars $\Xi_i(i = 0,\ldots,5)$ always exist such that

@@

\Vert H(q)\Vert & \ge \zeta_{min}(H(q))>\Xi_0>0\\
\Vert H(q)\Vert & \le \zeta_{max}(H(q))<\Xi_1<\infty\\
\Vert C(q,\dot q)\Vert & \le \Xi_2\Vert\dot q\Vert\\
\Vert G(q)\Vert & \le \Xi_3\\
\Vert \dot q_r\Vert & \le \Xi_4+\alpha\Vert\Delta q\Vert+\gamma\Vert\sigma\Vert\\
\Vert \ddot q_r\Vert & \le \Xi_5+\alpha\Vert\Delta \dot q\Vert\\

@@

where $\zeta_{min}(\cdot)$ and $\zeta_{max}(\cdot)$ mean the minimum and maximum eigenvalues of matrix, respectively, and $\alpha$ and $\gamma$ are controller parameters which will be discussed later on. Furthermore, one can have $Y_r\Theta\le \eta(t)$, in which $\eta(t)=f(\Delta q,\Delta \dot q,\sigma,\Xi_i,t)$ is a state-dependent function, thus $Y_r\Theta$ is usually thought to be a bounded function such that $\vert\eta\vert\le\varphi$, in which $\varphi$ is a positive scalar determined according to the system state~\cite{parra2003dynamic}.

@@

\overrightarrow{g}(\tau) &= \left(g(\hat\tau_t),g(\hat\tau_\theta),g(\hat\tau_\phi)\right)^T\\

@@

\overrightarrow{Sine}(q)=\begin{pmatrix}Sine(\lambda),Sine(\theta),Sine(\phi)\end{pmatrix}^T.

@@

\tau = -\overrightarrow{g}(\delta\varphi F(t) K_d\overrightarrow{sign}(S_{r1})),\label{eq:slider control}

@@

S_{q1} &= \overrightarrow{Sine}(\Delta \dot q+\alpha\Delta q),

@@

\frac{d}{dt}F(t) = \delta\varphi S_{r2}^T K_1K_de^{-2K_1^{-1}F^{-1}(t)}\overrightarrow{sign}(S_{r2})

@@

\overrightarrow{sign}(x) = \begin{pmatrix}sign(x_1),sign(x_2),\cdots,sign(x_n)\end{pmatrix}^T,

@@

for a vector $x=\left(x_1,x_2,,\cdots,x_n\right)^T$.

@@

S_{q2} &= \Delta \dot q+\alpha \Delta q,

@@

&S_{r2}^T \overrightarrow{sign}(S_{r2}) \ge  S_{r1}^T \overrightarrow{sign}(S_{r1})\\
&\overrightarrow{sign}(S_{r2})=\overrightarrow{sign}(S_{r1}),

@@

H(q)\dot S_{r2} = -\overrightarrow{g}(\delta\varphi F(t) K_d\overrightarrow{sign}(S_{r1}))-C(q,\dot q)S_{r2}-Y_{r2}\Theta.

@@

&= -S_{r2}^T\overrightarrow{g}(\delta\varphi F(t) K_d\overrightarrow{sign}(S_{r1}))-S_{r2}^TC(q,\dot q)S_{r2}+\frac{1}{2}S_{r2}^T\dot H(q)S_{r2}-S_{r2}^TY_{r2}\Theta+v\dot v\\
&= -S_{r2}^T\overrightarrow{g}(\delta\varphi F(t) K_d\overrightarrow{sign}(S_{r1}))-S_{r2}^TY_{r2}\Theta+\frac{1}{2}S_{r2}^T(\dot H(q)-2C(q,\dot q))S_{r2}+v\dot v\\
&= -S_{r2}^T\overrightarrow{g}(\delta\varphi F(t) K_d\overrightarrow{sign}(S_{r1}))-S_{r2}^TY_{r2}\Theta+v\dot v\\
&= -S_{r2}^TR_M(\delta\varphi F(t) K_d\overrightarrow{sign}(S_{r1}))\delta\varphi F(t) K_d\overrightarrow{sign}(S_{r1})-S_{r2}^TY_{r2}\Theta+v\dot v.\label{eq:V1_1}

@@

v\dot v =\varphi\delta S_{r2}^T K_1K_d(F(t)-K_1^{-1})\overrightarrow{sign}(S_{r2}),

@@

utilizing $\overrightarrow{sign}(S_{r1})=\overrightarrow{sign}(S_{r2})$ and $K_1$ proceeds to rewrite Eq.~(\ref{eq:V1_1}) as:

@@

\dot V&\le -S_{r2}^T\delta\varphi  K_1K_dF(t)\overrightarrow{sign}(S_{r2})+S_{r2}^T\varphi \overrightarrow{sign}(S_{r2})+\varphi\delta S_{r2}^T K_1K_d(F(t)-K_1^{-1})\overrightarrow{sign}(S_{r2})\\
&\le \varphi S_{r2}^T\overrightarrow{sign}(S_{r2})-\varphi\delta S_{r2}^T K_d\overrightarrow{sign}(S_{r2})

@@

\Delta \dot q=-\alpha \Delta q.

@@

Select the Lyapunov function candidate $V_2=\frac{1}{2}\Delta q^2$, the derivative of which follows

@@

\dot V_2 &= \Delta q\Delta \dot q\\
&=-\alpha (\Delta q)^2\le 0,

@@

if and only if $\Delta q = 0$, namely, $\Delta \dot q = 0$, one has $\dot V_2=0$. Hence asymptotically convergence of $\Delta q$ has been explained, namely, $\Delta q\rightarrow 0$ as $t\rightarrow \infty$. To conclude, the error state $\Delta q$ of the system~(\ref{eq:dynamics 1}) asymptotically converges to the origin with the limited input~(\ref{eq:slider control}). This completes the proof.

@@

\dot Z = \vert Z\vert^{\hat\alpha} \overrightarrow{sign}(Z),

@@

\tau = -\overrightarrow{g}(\delta\varphi F(t) K_{d1}\overrightarrow{sign}(S_{r3}))

@@

S_{r3} &= S_{q3}+\gamma\int^t_{t_0}\Vert S_{q3}(\varsigma)\Vert^{\hat\alpha}  \overrightarrow{sign}(S_{q3}(\varsigma))d\varsigma\label{eq:Sr3}\\
S_{q3} &= \Delta \dot q+\alpha \overrightarrow{Sine}(\xi\Delta q),\label{eq:Sq3}

@@

If the designer parameter $\hat\alpha = 1$, the quasi-saturated terminal sliding mode control degenerates into the quasi-saturated sliding mode control. Compared with the saturated sliding mode control, $S_{q3}$ converges to the origin in finite time in the quasi-saturated terminal method. This is because the establishment of the equation $\dot S_{q3} = \gamma\Vert S_{q3}\Vert^{\hat\alpha}  sign(S_{q3})$.

@@

\tau = -\overrightarrow{g}(\delta\varphi F(t) K_{d1}\overrightarrow{sign}(S_{r4})),

@@

S_{q4} &= \overrightarrow{Sine}(\Delta \dot q+\alpha\xi\Vert\Delta q\Vert^{\hat\alpha}\overrightarrow{sign}(\Delta q)),\label{eq:Sr4}

>>>>

H(\bm q)\ddot {\bm q}+C(\bm q,\dot{\bm q})\dot{\bm q}+G(\bm q) = \bm\tau,\label{eq:dynamic model}

@@

\bm q&= \begin{pmatrix}\lambda&\theta&\phi\end{pmatrix}^T\\
\bm \tau&=\begin{pmatrix}\hat\tau_t&\hat\tau_\theta&\hat\tau_\phi\end{pmatrix}^T\\
H(\bm q) &= \begin{bmatrix}1 &0&0\\0 &\lambda^2\cos^2\phi&0\\0&0&\lambda^2\end{bmatrix}\\
C(\bm q,\dot{\bm q}) &=\begin{bmatrix}0 &-(\lambda\dot \theta+2\lambda)\cos^2\phi&-\lambda\dot\phi\\(2\lambda+\lambda\dot \theta)\cos^2\phi&\lambda\dot \lambda\cos^2\phi-\lambda^2\dot\phi\sin\phi\cos\phi&-(\dot\theta+2)\lambda^2\sin\phi\cos\phi\\ \lambda\dot\phi&\lambda^2(\dot\theta+2)\sin\phi\cos\phi&\lambda\dot \lambda\end{bmatrix}\\
G(\bm q) &=\begin{bmatrix}-\lambda\cos^2\phi+2\lambda+3\lambda\cos ^2\theta\cos^2\phi\\3\lambda^2\cos\theta\sin\theta\cos^2\phi\\\lambda^2+3\lambda^2\cos^2\theta\sin\phi\cos\phi\end{bmatrix}.

@@

Based on the expression above, one can find that $\dot H(\bm q) - 2C(\bm q,\dot{\bm q})$ is skew-symmetric, namely,

@@

\bm q^T(\dot H(\bm q) - 2C(\bm q,\dot{\bm q}))\bm q = 0.

@@

Before declaring the controller design, some basic precondition is prepared for proceeding derivation.  The left side of Eq.~(\ref{eq:dynamic model}) is able to be denoted by the linearly parametrization $Y\Theta$, in which $Y=Y(\bm q,\dot {\bm q},\ddot {\bm q})$ consists of known nonlinear functions. $\Theta$ denotes unknown but constant parameters, and particularly it is a constant vector in this dimensionless form. Reconstruct the parametrization $Y\Theta$ in terms of a reference $\dot{\bm q}_r$:

@@

 Y_r\Theta\equiv H(\bm q)\ddot {\bm q}_r+C(\bm q,\dot {\bm q})\dot {\bm q}_r+G(\bm q)\label{eq:dynamic model 1}

@@

where $Y_r = Y_r(\bm q,\dot{\bm q},\dot{\bm q}_r,\ddot{\bm q}_r)$. Based on results mentioned above, the open-loop error dynamics expressed by $S_r$ is able to be obtained, as follows:

@@

H(\bm q)\dot S_r+C(\bm q,\dot{\bm q})S_r=\bm\tau - Y_r\Theta\label{eq:dynamic model 2}

@@

S_r=\dot {\bm q}-\dot {\bm q}_r.\label{eq:Sr1}

@@

We define the tracking error as $\Delta \bm q=\bm q-\bm q_d$, in which $\bm q_d$ is the reference trajectory with $\bm q_d(t)\in C^2$. For deriving the control scheme, some properties should be given out in advance.
\begin{mypro}\cite{parra2003dynamic}
Positive scalars $\Xi_i(i = 0,\ldots,6)$ always exist such that

@@

\Vert H(\bm q)\Vert & \ge \zeta_{min}(H(\bm q))>\Xi_0>0\\
\Vert H(\bm q)\Vert & \le \zeta_{max}(H(\bm q))<\Xi_1<\infty\\
\Vert C(\bm q,\dot {\bm q})\Vert & \le \Xi_2\Vert\dot {\bm q}\Vert+\Xi_3\\
\Vert G(\bm q)\Vert & \le \Xi_4\\
\Vert \dot {\bm q}_r\Vert & \le \Xi_5+\alpha\Vert\Delta \bm q\Vert+\gamma\Vert\sigma\Vert\\
\Vert \ddot {\bm q}_r\Vert & \le \Xi_6+\alpha\Vert\Delta \dot {\bm q}\Vert\\

@@

where $\zeta_{min}(\cdot)$ and $\zeta_{max}(\cdot)$ mean the minimum and maximum eigenvalues of matrix, respectively, and $\alpha$ and $\gamma$ are controller parameters which will be discussed later on. Furthermore, one can have $Y_r\Theta\le \eta(t)$, in which $\eta(t)=f(\Delta \bm q,\Delta \dot {\bm q},\sigma,\Xi_i,t)$ is a state-dependent function, thus $Y_r\Theta$ is usually thought to be a bounded function such that $\vert\eta\vert\le\varphi$, in which $\varphi$ is a positive scalar determined according to the system state.

@@

\bm{g}(\tau) &= \left(g(\hat\tau_t),g(\hat\tau_\theta),g(\hat\tau_\phi)\right)^T\\

@@

\bm{Sine}(\bm q)=\begin{pmatrix}Sine(\lambda),Sine(\theta),Sine(\phi)\end{pmatrix}^T.

@@

\bm\tau = -\bm{g}(\delta\varphi F(t) K_d\bm{sign}(S_{r1})),\label{eq:slider control}

@@

S_{q1} &= \bm{Sine}(\Delta \dot {\bm q}+\alpha\Delta\bm q),

@@

\frac{d}{dt}F(t) = \delta\varphi S_{r2}^T K_1K_de^{-2K_1^{-1}F^{-1}(t)}\bm{sign}(S_{r2})

@@

\bm{sign}(\bm x) = \left(sign(x_1),sign(x_2),\cdots,sign(x_n)\right)^T,

@@

for a vector $\bm x=\left(x_1,x_2,,\cdots,x_n\right)^T$.

@@

S_{q2} &= \Delta \dot {\bm q}+\alpha \Delta \bm q,

@@

&S_{r2}^T \bm{sign}(S_{r2}) \ge  S_{r1}^T \bm{sign}(S_{r1})\\
&\bm{sign}(S_{r2})=\bm{sign}(S_{r1}),

@@

H(q)\dot S_{r2} = -\bm{g}(\delta\varphi F(t) K_d\bm{sign}(S_{r1}))-C(q,\dot q)S_{r2}-Y_{r2}\Theta.

@@

&= -S_{r2}^T\bm{g}(\delta\varphi F(t) K_d\bm{sign}(S_{r1}))-S_{r2}^TC(q,\dot q)S_{r2}+\frac{1}{2}S_{r2}^T\dot H(q)S_{r2}-S_{r2}^TY_{r2}\Theta+v\dot v\\
&= -S_{r2}^T\bm{g}(\delta\varphi F(t) K_d\bm{sign}(S_{r1}))-S_{r2}^TY_{r2}\Theta+\frac{1}{2}S_{r2}^T(\dot H(q)-2C(q,\dot q))S_{r2}+v\dot v\\
&= -S_{r2}^T\bm{g}(\delta\varphi F(t) K_d\bm{sign}(S_{r1}))-S_{r2}^TY_{r2}\Theta+v\dot v\\
&= -S_{r2}^TR_M(\delta\varphi F(t) K_d\bm{sign}(S_{r1}))\delta\varphi F(t) K_d\bm{sign}(S_{r1})-S_{r2}^TY_{r2}\Theta+v\dot v.\label{eq:V1_1}

@@

v\dot v =\varphi\delta S_{r2}^T K_1K_d(F(t)-K_1^{-1})\bm{sign}(S_{r2}),

@@

utilizing $\bm{sign}(S_{r1})=\bm{sign}(S_{r2})$ and $K_1$ proceeds to rewrite Eq.~(\ref{eq:V1_1}) as:

@@

\dot V&\le -S_{r2}^T\delta\varphi  K_1K_dF(t)\bm{sign}(S_{r2})+S_{r2}^T\varphi \bm{sign}(S_{r2})+\varphi\delta S_{r2}^T K_1K_d(F(t)-K_1^{-1})\bm{sign}(S_{r2})\\
&\le \varphi S_{r2}^T\bm{sign}(S_{r2})-\varphi\delta S_{r2}^T K_d\bm{sign}(S_{r2})

@@

\Delta \dot {\bm q}=-\alpha \Delta \bm q.

@@

Select the Lyapunov function candidate $V_2=\frac{1}{2}\Delta \bm q^T\Delta\bm q$, the derivative of which follows

@@

\dot V_2 &= \Delta\bm q^T\Delta \dot{\bm q}\\
&=-\alpha \Delta \bm q^T\Delta \bm q\le 0,

@@

if and only if $\Delta \bm q = 0$, namely, $\Delta \dot {\bm q} = 0$, one has $\dot V_2=0$. Hence asymptotically convergence of $\Delta \bm q$ has been explained, namely, $\Delta \bm q\rightarrow 0$ as $t\rightarrow \infty$. To conclude, the error state $\Delta \bm q$ of the system~(\ref{eq:dynamics 1}) asymptotically converges to the origin with the limited input~(\ref{eq:slider control}). This completes the proof.

@@

\dot Z = \vert Z\vert^{\hat\alpha} \bm{sign}(Z),

@@

\tau = -\bm{g}(\delta\varphi F(t) K_{d1}\bm{sign}(S_{r3}))

@@

S_{r3} &= S_{q3}+\gamma\int^t_{t_0}\Vert S_{q3}(\varsigma)\Vert^{\hat\alpha}  \bm{sign}(S_{q3}(\varsigma))d\varsigma\label{eq:Sr3}\\
S_{q3} &= \Delta \dot q+\alpha \bm{Sine}(\xi\Delta q),\label{eq:Sq3}

@@

If the designer parameter $\hat\alpha = 1$, the quasi-saturated terminal sliding mode control degenerates into the quasi-saturated sliding mode control. Compared with the saturated sliding mode control, $S_{q3}$ converges to the origin in finite time in the quasi-saturated terminal method. This is because the establishment of the equation $\dot S_{q3} = \gamma\Vert S_{q3}\Vert^{\hat\alpha} \bm{ sign}(S_{q3})$.

@@

\tau = -\bm{g}(\delta\varphi F(t) K_{d1}\bm{sign}(S_{r4})),

@@

S_{q4} &= \bm{Sine}(\Delta \dot q+\alpha\xi\Vert\Delta q\Vert^{\hat\alpha}\bm{sign}(\Delta q)),\label{eq:Sr4}
